---
title: "Playing with Oceanographic Data"
author: "Thomas Bryce Kelly"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gridding Datasets with TheSource}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(TheSource)
```

```{r}
## Load in the data from BCO-DMO
# Note: the unit line is the second row, so to get the header values and the data we have to use two read operations.
data = read.csv('https://erddap.bco-dmo.org/erddap/tabledap/bcodmo_dataset_668083.csv', stringsAsFactors = F, skip = 2, header = F)
colnames(data) = read.csv('https://erddap.bco-dmo.org/erddap/tabledap/bcodmo_dataset_668083.csv', stringsAsFactors = F, nrows = 1, header = F)

```


First, we will plot a map of the stations so we can make sense of where they are.
```{r}
map = make.map(lon.min = min(data$longitude), lon.max = max(data$longitude),
               lat.min = min(data$latitude), lat.max = max(data$latitude))
add.map.text(data$longitude, data$lat, data$STNNBR)
```

Let's plot a section using stations >=7
```{r}
l = which(data$STNNBR >= 7)

section = build.section(x = data$longitude[l],
                        y = data$DEPTH_MAX[l],
                        z = data$Y_SPT_CONC_PUMP[l],
                        lat = data$latitude[l],
                        lon = data$longitude[l], field.names = 'Yttrium', x.factor = 50)

plot.section(section, mark.points = T, ylim = c(5000, 0), zlim = c(0,4))
add.section.bathy(section, bathy.col = '#555555e0')

```

How about a higher resolution interpolation?

```{r}
l = which(data$STNNBR >= 7)

section = build.section(x = data$longitude[l],
                        y = data$DEPTH_MAX[l],
                        z = data$Y_SPT_CONC_PUMP[l],
                        lat = data$latitude[l],
                        lon = data$longitude[l], field.names = 'Yttrium',
                        x.scale = 0.2, y.scale = 10,
                        x.factor = 50)

plot.section(section, mark.points = T, ylim = c(5000, 0), zlim = c(0, 4))
add.section.bathy(section, bathy.col = '#555555e0')

```

And let's add a color bar:

```{r}
par(plt = c(0.1, 0.8, 0.1, 0.9))
plot.section(section, mark.points = T, ylim = c(5000, 0), zlim = c(0, 4))
add.section.bathy(section, bathy.col = '#555555e0')
add.colorbar(min = 0, max = 4, labels = c(1:4), pal = ocean.matter)
```


```{r}
l = which(data$STNNBR >= 7)

section.idw = build.section(x = data$longitude[l],
                        y = data$DEPTH_MAX[l],
                        z = data$Y_SPT_CONC_PUMP[l],
                        lat = data$latitude[l],
                        lon = data$longitude[l], field.names = 'Yttrium',
                        nx = 100, ny = 100,
                        x.factor = 50)

section.NN = build.section(x = data$longitude[l],
                        y = data$DEPTH_MAX[l],
                        z = data$Y_SPT_CONC_PUMP[l],
                        lat = data$latitude[l],
                        lon = data$longitude[l], field.names = 'Yttrium',
                        nx = 100, ny = 100,
                        x.factor = 50, gridder = gridNN)

section.NNI = build.section(x = data$longitude[l],
                        y = data$DEPTH_MAX[l],
                        z = data$Y_SPT_CONC_PUMP[l],
                        lat = data$latitude[l],
                        lon = data$longitude[l], field.names = 'Yttrium',
                        nx = 100, ny = 100,
                        x.factor = 50, gridder = gridNNI)

section.krige = build.section(x = data$longitude[l],
                        y = data$DEPTH_MAX[l],
                        z = data$Y_SPT_CONC_PUMP[l],
                        lat = data$latitude[l],
                        lon = data$longitude[l], field.names = 'Yttrium',
                        nx = 100, ny = 100,
                        x.factor = 50, gridder = gridKrig)



```

Let's take a look at how the various gridding products compare graphically. The __kriging__ product does a good job of minimizing the influence of the concentration peak around 1500m. Unfortunately, this peak also corresponds to a low-sampling resolution region in the data overall as seen in the __nearest neighbor__ interpolation (the peak corresponds to a large polygon rather than a small one). This has a direct impact on the __natural neighbor__ interpolation. The __inverse distance__ interpolation falls somewhere between these two more extreme treatments of that peak with smooth transitions and a more limited zone of influence (compared to the __NNI__). 

```{R}

## Standard IDW
plot.section(section.idw, mark.points = T, ylim = c(5000, 0), main = 'IDW', zlim = c(0, 4))
add.section.bathy(section.idw, bathy.col = '#555555e0')

## NN
plot.section(section.NN, mark.points = T, ylim = c(5000, 0), main = 'Nearest Neighbor', zlim = c(0, 4))
add.section.bathy(section.NN, bathy.col = '#555555e0')

## NNI
plot.section(section.NNI, mark.points = T, ylim = c(5000, 0), main = 'Natural Neighbor', zlim = c(0, 4))
add.section.bathy(section.NNI, bathy.col = '#555555e0')

## Krige
plot.section(section.krige, mark.points = T, ylim = c(5000, 0), main = 'Kriging', zlim = c(0, 4))
add.section.bathy(section.krige, bathy.col = '#555555e0')

```

Since the sections were plotted with the same grid constraints (extent and resolution), we can simply subtract the (log) values to generate a new section. We can then plot this using the name we provide:

```{R}
section.idw$grid$NNI.minus.IDW = section.NNI$grid$Yttrium - section.idw$grid$Yttrium

plot.section(section.idw, field = 'NNI.minus.IDW', ylim = c(5000, 0), zlim = c(-2, 2), pal = ocean.delta)
add.section.bathy(section.idw, bathy.col = '#555555e0')

```



